---
title: "Experiment 2"
author: "Christopher M. Federico"
date: "08/25/2023"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options: 
  markdown: 
    wrap: 72
---

## Results for Experiment 2

Experiment 2 reverses (and expands) the logic of Experiment 1.
Experiment 2 replicates the finding that the effect of authoritarianism
on GOP preference is reduced when respondents are faced with a choice
where one of the candidates takes a party-incongruent stand on a social
issue. The baseline again presents a combination of background
biographical information and partisanship.

With this baseline, the idea is to establish the relationship between
authoritarianism and candidate preference that exists on the basis of
partisan cues alone. There are four treatment conditions that present
the same party and biographical information, but also differing
information about the signature issues of the candidates. The positions
taken by the Democratic candidate are always be party-congruent (with
the issue position varying across treatments), while the positions taken
by the Republican candidate are always party incongruent (with the issue
position differing across conditions):

-   control: party only
-   treatment 1: party + econ liberal D, social liberal R (immigration)
-   treatment 2: party + econ liberal D, social liberal R (LGBT/gender)
-   treatment 3: party + social liberal D (immigration), econ liberal R
-   treatment 4: party + social liberal D (LGBT/gender), econ liberal R

If things work as they did in Experiment 1, then we should expect a
positive relationship between authoritarianism and GOP preference in the
control and in any treatment where the Republican candidate DOES NOT
take a socially-liberal position. In contrast, the relationship between
authoritarianism and GOP preference should be weakened relative to the
control in treatments where the Republican candidate DOES take a
socially-liberal positions.

So, we would expect roughly similar positive relationships between
authoritarianism and GOP preference in the control, treatment 3, and
treatment 4, and a null or reduced-magnitude relationship between
authoritarianism and GOP preference (relative to control) in treatment 1
and treatment 2.

Beyond these basic predictions, Experiment 2 builds in an exploratory
element. Specifically, comparison between treatments 2 and 3 will allow
us to see whether immigration or LGBT issue defection is more
off-putting to authoritarians (or whether they are similarly
off-putting).

Load in preliminaries and coded data:

```{r message=FALSE}

#### packages.
library(car)
library(psych)
library(ggplot2)
library(sandwich)
library(expss)
library(summarytools)
library(tidyverse)
library(dotwhisker)
library(lmtest)
library(huxtable)
library(haven)
library(interactions)

################################################################################
######### utility functions

std01<-function(x){
  min.x<-min(x, na.rm=T)
  max.x<-max(x-min.x, na.rm=T)
  return((x-min.x)/max.x)
}
rsumm<-function(x){
  descr(x, stats='common', transpose=T)
}
tab<-function(x){
  freq(as_factor(x))
}

################################################################################
######### persistent Weber ggplot() theme

ggtheme = theme(
  plot.title=element_text(face="bold",hjust=0,vjust=0,colour="#3C3C3C",size=20),
  axis.text.x=element_text(size=16,colour="#535353",face="bold"),
  axis.text.y=element_text(size=16,colour="#535353",face="bold"),
  axis.title = element_text(size=16,colour="#535353",face="bold"),
  axis.title.y=element_text(size=16,colour="#535353",face="bold",vjust=1.5),
  axis.ticks=element_blank(),
  strip.text.x = element_text(size = 16),
  panel.grid.major=element_line(colour="#D0D0D0",size=.25),
  panel.background=element_rect(fill="white"),
  legend.text=element_text(size=14),
  legend.title=element_text(size=16)) 

################################################################################

##### load Rdata if coded and saved previously
load("wff2.Rdata") 

```

Subset to white respondents who passed manipulation checks (N=1140):

```{r message=FALSE}

###### subset to white respondents
wdata<-subset(data, white==1 & bc1==1 & sc1==1)

```

Worth nothing that authoritarianism forms a reliable scale:

```{r message=FALSE}

psych::alpha(with(wdata, cbind(aut1, aut2, aut3, aut4, aut5, aut6, aut7, aut8)))

```

## Vote choice model fit and conditional effects

Here is the setup of the conditions:

-   control = party only
-   treat1 = party + social lib R, econ lib D (immig)
-   treat2 = party + social lib R, econ lib D (LGBT)
-   treat3 = party + econ lib R, social lib D (immig)
-   treat4 = party + econ lib R, social lib D (LGBT)

Here, we see more or less what we expected. There is a clear positive
relation between authoritarianism and Republican vote preference in the
control condition, and this remains more or less the same in the
conditions where the Republican is NOT socially liberal. However, the
effect of authoritarianism drops in the conditions where the Republican
is socially liberal:

```{r message=FALSE}

# fit
m1 <- lm(rvc ~ female+age+rinc+college+jewish+cath+other+raut*cond, data=wdata)
coeftest(m1, vcovHC(m1, type = "HC3"))

# conditional effects
sim_margins(m1, pred = raut, modx = cond, robust=T) 

```

Plotting this:

```{r message=FALSE, fig.width=11, fig.height=8}

interact_plot(m1, pred = raut, 
                    modx = cond,
                    modx.labels=c("Party Only", 
                                  "Party + Soc Lib R, Econ Lib D\n(Immigration)", 
                                  "Party + Soc Lib R, Econ Lib D\n(LGBTQ)",
                                  "Party + Econ Lib R, Soc Lib D\n(Immigration)",
                                  "Party + Econ Lib R, Soc Lib D\n(LGBTQ)"), 
                    interval = TRUE,
                    robust=T, 
                    legend.main = "Condition",
                    data = wdata) + 
  ggtheme +
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1)) + 
  scale_x_continuous(breaks=seq(0,1,0.2)) + 
  theme(aspect.ratio=1, 
        plot.title = element_text(hjust = 0.5)) +  
  labs(title = "Candidate Support by Condition (Linear)",
       x = "Authoritarianism",
       y = "Support for Republican Candidate")

```

## Feeling thermometer difference model fit and conditional effects

The second analysis used the Republican candidate FT minus the
Democratic candidate FT (with the difference score recoded 0-1) as the
dependent measure.

The results are more or less the same as above, but a bit more clearly
consistent with our predictions:

```{r message=FALSE}

# fit
m2 <- lm(ftdif ~ female+age+rinc+college+jewish+cath+other+raut*cond, 
         data=wdata)
coeftest(m2, vcovHC(m2, type = "HC3"))

# conditional effects
sim_margins(m2, pred = raut, modx = cond, robust=T) 

```

Plotting this:

```{r message=FALSE, fig.width=11, fig.height=8}

interact_plot(m2, pred = raut, 
                    modx = cond,
                    modx.labels=c("Party Only", 
                                  "Party + Soc Lib R, Econ Lib D\n(Immigration)", 
                                  "Party + Soc Lib R, Econ Lib D\n(LGBTQ)",
                                  "Party + Econ Lib R, Soc Lib D\n(Immigration)",
                                  "Party + Econ Lib R, Soc Lib D\n(LGBTQ)"), 
                    interval = TRUE,
                    robust=T, 
                    legend.main = "Condition",
                    data = wdata) + 
  ggtheme +
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1)) + 
  scale_x_continuous(breaks=seq(0,1,0.2)) + 
  theme(aspect.ratio=1, 
        plot.title = element_text(hjust = 0.5)) +  
  labs(title = "Comparative Candidate FT Ratings\nby Condition (Linear)",
       x = "Authoritarianism",
       y = "Evaluate R More Positively than D")

```

## Candidate trait attribution difference model fit and conditional effects

The third analysis averaged respondents' ratings of each candidate on
effectiveness and shared values. Then the Republican candidate average
minus the Democratic candidate average was computed (with the difference
score recoded 0-1). This served as the dependent measure.

Again, we get the expected result:

```{r message=FALSE}

# fit
m3 <- lm(bsdif2 ~ female+age+rinc+college+jewish+cath+other+raut*cond, 
         data=wdata)
coeftest(m3, vcovHC(m3, type = "HC3"))

# conditional effects
sim_margins(m3, pred = raut, modx = cond, robust=T) 

```

Plotting this:

```{r message=FALSE, fig.width=11, fig.height=8}

interact_plot(m3, pred = raut, 
                    modx = cond,
                    modx.labels=c("Party Only", 
                                  "Party + Soc Lib R, Econ Lib D\n(Immigration)", 
                                  "Party + Soc Lib R, Econ Lib D\n(LGBTQ)",
                                  "Party + Econ Lib R, Soc Lib D\n(Immigration)",
                                  "Party + Econ Lib R, Soc Lib D\n(LGBTQ)"), 
                    interval = TRUE,
                    robust=T, 
                    legend.main = "Condition",
                    data = wdata) + 
  ggtheme +
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1)) + 
  scale_x_continuous(breaks=seq(0,1,0.2)) + 
  theme(aspect.ratio=1, 
        plot.title = element_text(hjust = 0.5)) +  
  labs(title = "Comparative Candidate Traits\nby Condition (Linear)",
       x = "Authoritarianism",
       y = "Attribute More Positive Traits to R than D")

```

So, for the most part, we find what we expected in the full version of the
experiment.

## Collapsing the Treatments

Now, the differences between the LGBT and immigration versions of the treatments are not terribly profound, so I also collapsed conditions 1+2 and conditions 3+4. This gives us the following simplified setup:

-   control = party only
-   treat1 = party + social lib R, econ lib D
-   treat2 = party + econ lib R, social lib D

The results are somewhat clearer in the collapsed setup, as shown below. (The variable for collapsed condition is 'condc'.)

## Vote choice model fit and conditional effects (collapsed)

Here, we see more or less what we expected. There is a clear positive
relation between authoritarianism and Republican vote preference in the
control condition, and this remains more or less the same in the
collapsed condition where the Republican is NOT socially liberal.
However, the effect of authoritarianism drops in the collapsed condition
where the Republican is socially liberal:

```{r message=FALSE}

# fit
m1c <- lm(rvc ~ female+age+rinc+college+jewish+cath+other+raut*condc, data=wdata)
coeftest(m1c, vcovHC(m1c, type = "HC3"))

# conditional effects
sim_margins(m1c, pred = raut, modx = condc, robust=T) 

```

Plotting this:

```{r message=FALSE, fig.width=11, fig.height=8}

interact_plot(m1c, pred = raut, 
                     modx = condc,
                     modx.labels=c("Party Only", 
                                   "Party + Soc Lib R, Econ Lib D", 
                                   "Party + Econ Lib R, Soc Lib D"), 
                     interval = TRUE,
                     robust=T, 
                     legend.main = "Condition",
                     data = wdata) + 
  ggtheme +
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1)) + 
  scale_x_continuous(breaks=seq(0,1,0.2)) + 
  theme(aspect.ratio=1, 
        plot.title = element_text(hjust = 0.5)) +  
  labs(title = "Candidate Support by Condition (Linear)",
       x = "Authoritarianism",
       y = "Support for Republican Candidate")

```

## Feeling thermometer difference model fit and conditional effects (collapsed)

The second analysis used the Republican candidate FT minus the
Democratic candidate FT (with the difference score recoded 0-1) as the
dependent measure.

The results are more or less the same as above:

```{r message=FALSE}

# fit
m2c <- lm(ftdif ~ female+age+rinc+college+jewish+cath+other+raut*condc, 
         data=wdata)
coeftest(m2c, vcovHC(m2c, type = "HC3"))

# conditional effects
sim_margins(m2c, pred = raut, modx = condc, robust=T) 

```

Plotting this:

```{r message=FALSE, fig.width=11, fig.height=8}

interact_plot(m2c, pred = raut, 
                     modx = condc,
                     modx.labels=c("Party Only", 
                                   "Party + Soc Lib R, Econ Lib D", 
                                   "Party + Econ Lib R, Soc Lib D"), 
                     interval = TRUE,
                     robust=T, 
                     legend.main = "Condition",
                     data = wdata) + 
  ggtheme +
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1)) + 
  scale_x_continuous(breaks=seq(0,1,0.2)) + 
  theme(aspect.ratio=1, 
        plot.title = element_text(hjust = 0.5)) +  
  labs(title = "Comparative Candidate FT Ratings\nby Condition (Linear)",
       x = "Authoritarianism",
       y = "Evaluate R More Positively than D")

```

## Candidate trait attribution difference model fit and conditional effects (collapsed)

The third analysis averaged respondents' ratings of each candidate on
effectiveness and shared values. Then the Republican candidate average
minus the Democratic candidate average was computed (with the difference
score recoded 0-1). This served as the dependent measure.

As before, we get the expected result:

```{r message=FALSE}

# fit
m3c <- lm(bsdif2 ~ female+age+rinc+college+jewish+cath+other+raut*condc, 
         data=wdata)
coeftest(m3c, vcovHC(m3c, type = "HC3"))

# conditional effects
sim_margins(m3c, pred = raut, modx = condc, robust=T) 

```

Plotting this:

```{r message=FALSE, fig.width=11, fig.height=8}

interact_plot(m3c, pred = raut, 
                     modx = condc,
                     modx.labels=c("Party Only", 
                                   "Party + Soc Lib R, Econ Lib D", 
                                   "Party + Econ Lib R, Soc Lib D"), 
                     interval = TRUE,
                     robust=T, 
                     legend.main = "Condition",
                     data = wdata) + 
  ggtheme +
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1)) + 
  scale_x_continuous(breaks=seq(0,1,0.2)) + 
  theme(aspect.ratio=1, 
        plot.title = element_text(hjust = 0.5)) +  
  labs(title = "Comparative Candidate Traits\nby Condition (Linear)",
       x = "Authoritarianism",
       y = "Attribute More Positive Traits to R than D")

```

## In Sum

All in all, Experiment 2 produced results that were consistent with our
predictions and which conceptually replicated the reversed design we
used in Experiment 1. I think these two experiments would form a good
core for an experimental chapter that we could put right after the panel
data chapter.
